Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-30 paper/abstract.tex
General pass. Paper still needs compressing

  % Status
  Automatic generation of random values described by abstract data types (ADTs)
  is often a hard task.
  % data types (ADTs), % ---specially when considering ADTs which
  % represent HTML pages, images, programs or other structural rich objects.
  % Situation
  State-of-the-art random testing tools often can automatically synthesize
  random data generators based on ADTs definitions.
  %
  In that manner, generated values comply with the structure described by
  ADTs---something that proves useful when testing software which expects
  complex input like web pages, images, or even programs.
  %
  % Problem
  However, it sometimes becomes necessary to generate structural richer ADTs
  values in order to test deeper software layers.
  %
  % Solution
  In this light, we propose to leverage static information found in the
  codebase as a manner to improve the generation process.
  %
  Namely, our generators are capable to consider how programs branch on input
  data as well as how ADTs values are built via interfaces.
  %
  % Concretely, we propose to consider not only ADTs definitions, but also how
  % programs branch on input data as well as how ADTs values get manipulated via
  % interfaces.
  %
  We implement a tool, called {\dragenp}, responsible to synthezise generators
  for ADTs values while providing compile-time guarantees about their
  distributions.
  %
  Our compile-time predictions allow {\dragenp} to provide an heuristic that
  tries to adjust the distribution of generators to what developers might want.
  \todo[inline,author=Ale]{Case study}

\end{abstract}

>>>>

  % Status
  Automatic generation of random values described by algebraic data types (ADTs)
  is often a hard task.
  % data types (ADTs), % ---specially when considering ADTs which
  % represent HTML pages, images, programs or other structural rich objects.
  % Situation
  State-of-the-art random testing tools can automatically synthesize random data
  generators based on ADTs definitions.
  %
  % In that manner, generated values comply with the structure described by
  % ADTs---something that proves useful when testing software which expects
  % complex input like web pages, images, or even programs.
  %
  In that manner, generated values comply with the structure described by ADTs,
  something that proves useful when testing software which expects complex
  inputs.
  %
  % Problem
  However, it sometimes becomes necessary to generate structural richer ADTs
  values in order to test deeper software layers.
  %
  % Solution
  % In this light, we propose to leverage static information found in the
  In this work we propose to leverage static information found in the codebase
  as a manner to improve the generation process.
  %
  Namely, our generators are capable to consider how programs branch on input
  data as well as how ADTs values are built via interfaces.
  %
  % Concretely, we propose to consider not only ADTs definitions, but also how
  % programs branch on input data as well as how ADTs values get manipulated via
  % interfaces.
  %
  We implement a tool, responsible to synthezise generators for ADTs values
  while providing compile-time guarantees about their distributions.
  %
  Using compile-time predictions, we provide an heuristic that tries to adjust
  the distribution of generators to what developers might want.
  % \todo[inline,author=Ale]{Case study}
  Preliminary experiments of our approach are very encouraging.
\end{abstract}

### 2019-01-30 paper/casestudies.tex
General pass. Paper still needs compressing

\section{Case Studies}

@@

\begin{figure*}[t]
  \centering
  \input{tikz/lisp.tex}
  \hspace{5pt}%
  \input{tikz/html.tex}
  \caption{Path coverage comparison between \dragen (\ref{exp:dragen}) and
    \dragenp (\ref{exp:dragenp}). }
  \label{fig:coverage}
\end{figure*}
%
Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
triggered for different sizes of the corpus size and derivation tool, including
error bars indicating the standard error of the mean on each case.
%
We proceed to describe each case study and our findings in detail as follows.

@@

Our chosen library encodes Lisp S-expressions essentially as lists of symbols,
represented as plain strings; and literal values like booleans or integers.
%
In order to interpret Lisp programs, this unified representation of data and
code requires this library to pattern match against common patterns like
let-bindings, if-then-else expressions and arithmetic operators among others.
%
In particular, each one of these patterns match a against special symbol of the
Lisp syntax like |"let"|, |"if"| or |"+"|; and their corresponding
sub-expressions.
%
We extracted this structural information and included it into the generation
specification of our random Lisp values---which were generated by randomly
picking from a total of 6 data constructors and 8 different pattern matchings.
%
By doing this, we obtained a code coverage improvement of approximately $4\%$
using \dragenp with respect to the one obtained with \dragen (see Figure
\ref{fig:coverage} (a)).
%
While it seems a little improvement, it is worth noticing that we are generating
highly structured input, i.e., Lisp programs.
%
When testing interpreters, we argue that an improvement of $4\%$ is not
negligible considering the little effort that took us to specify the pattern
matchings.
%\todo[inline, author=AM]{Shall we say anything justifing this little
%  improvement?}

>>>>

\section{Case Studies} \label{sec:casestudies}

@@

\begin{figure*}[t]
  \centering
  \input{tikz/lisp.tex}
  \hspace{5pt}%
  \input{tikz/html.tex}
  \caption{Path coverage comparison between \dragen (\ref{exp:dragen}) and
    \dragenp (\ref{exp:dragenp}). }
  \label{fig:coverage}
\end{figure*}
%
Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
triggered for different combinations of corpus size and derivation tool,
including error bars indicating the standard error of the mean on each case.
%
We proceed to describe each case study and our findings in detail as follows.

@@

Our chosen library encodes Lisp S-expressions essentially as lists of symbols,
represented as plain strings; and literal values like booleans or integers.
%
In order to interpret Lisp programs, this unified representation of data and
code requires this library to pattern match against common patterns like
let-bindings, if-then-else expressions and arithmetic operators among others.
%
In particular, each one of these patterns match a against special symbol of the
Lisp syntax like |"let"|, |"if"| or |"+"|; and their corresponding
sub-expressions.
%
We extracted this structural information and included it into the generation
specification of our random Lisp values---which were generated by randomly
picking from a total of 6 data constructors and 8 different pattern matchings.
%
By doing this, we obtained a code coverage improvement of approximately $4\%$
using \dragenp with respect to the one obtained with \dragen (see Figure
\ref{fig:coverage} (a)).
%
While it seems a little improvement, it is worth noticing that we are generating
highly structured inputs, i.e., Lisp programs.
%
When testing interpreters, we argue that an improvement of $4\%$ is not
negligible considering the little effort that took us to specify the pattern
matchings.
%\todo[inline, author=AM]{Shall we say anything justifing this little
%  improvement?}

### 2019-01-30 paper/introduction.tex
General pass. Paper still needs compressing

% Scenario of random testing
Random testing is a promising approach to finding bugs
\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
%
\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
Haskell community.
%
It requires developers to specify (i) \emph{testing properties} which describe
programs behavior and (ii) \emph{random data generator} based on the
{\emph{types}} of the expected inputs (e.g., an integer, an string,
etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
%
QuickCheck then generates random test cases and reports those violating testing
properties.
%

% Situation with Quick Check, ADTs, tools
QuickCheck comes equipped with random generators for built-in types, while it
requires to manually write generators for user-defined ADTs.
%
Recently, there have been a proliferation of tools to automatically derive
QuickCheck generators for ADTs
\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
%
The main difference about these tools lies on the guarantees provided to ensure
\emph{the termination of the generation process} and the \emph{distribution of
  random values}.
%
Despite their differences, these tools guarantee that generated values are
\emph{well-typed}.
%
In other words, generated values follow the structure described by the
definition of the ADT.
%

%% How to use random generated ADTs
Well-typed ADT values are specially useful when testing programs which expect
highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
%
Generating ADT values also proves fruitful when looking for vulnerabilities with
fuzzers \cite{GriecoCB16,grieco2017}.
%
%% Problem
%
Despite these success stories, ADT type-definitions do not often capture all the
invariants expected from the data that they are intended to model.
%
As a result, even if random values are well-typed, they might not be built with
enough structure to penetrate into deep layers of software.

%% Our proposal
In this work, we propose a novel improvement in the generation process of ADT
values by exploiting some static information found in the codebase.
%
More specifically, to refine the structure of generated values, we propose a
generation process that is capable to consider how programs branch based on ADTs
values' patterns as well as how they get manipulated by abstract interfaces.
%
Furthermore, we show how to predict the distribution of the \emph{expected}
numbers of ADT constructors, values fitting certain branching patterns, and
calls to interfaces.
%
For that, we extend some recent results on applying \emph{branching
  processes}\cite{gw1875}---a simple stochastic model conceived to study
population growth.
%---to predict the distribution of QuickCheck
%generators\cite{DBLP:conf/haskell/MistaRH18}.
%
We implement our ideas in a tool, called {\dragenp}, that is capable to
automatically synthesize QuickCheck generators for \emph{rich ADT values}, where
the distributions of random values can be adjusted at compile-time to what
developers might want based on our predictions.
%
\todo[inline,author=Ale]{Add here later about test cases when we know what they
  are}

We remark that, although this work focuses on Haskell algebraic data types, this
technique is general enough to be applied to most programming languages.
%with some level of support for composite types.

>>>>

% Scenario of random testing
Random testing is a promising approach to finding bugs
\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
%
\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
Haskell community.
%
It requires developers to specify (i) \emph{testing properties} describing
programs expected behavior and (ii) \emph{random data generators} based on the
\emph{types} of the expected inputs (e.g., an integer, an string,
etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
%
\quickcheck then generates random test cases and reports those violating testing
properties.

% Situation with Quick Check, ADTs, tools
\quickcheck comes equipped with random generators for built-in types, while it
requires to manually write generators for user-defined ADTs.
%
Recently, there have been a proliferation of tools to automatically derive
\quickcheck generators for ADTs
\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
%
The main difference about these tools lies on the guarantees provided to ensure
\emph{the termination of the generation process} and the \emph{distribution of
  random values}.
%
Despite their differences, these tools guarantee that generated values are
\emph{well-typed}.
%
In other words, generated values follow the structure described by the
definition of the ADT.

%% How to use random generated ADTs
Well-typed ADT values are specially useful when testing programs which expect
highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
%
Generating ADT values also proves fruitful when looking for vulnerabilities in
combination with fuzzers \cite{GriecoCB16,grieco2017}.
%
%% Problem
%
Despite these success stories, ADT type-definitions do not often capture all the
invariants expected from the data that they are intended to model.
%
As a result, even if random values are well-typed, they might not be built with
enough structure to penetrate into deep layers of software.

%% Our proposal
In this work, we identify two different sources of structural information that
can be statically exploited to improve the generation process of ADT values
(Section \ref{sec:sources}).
%
% In this work, we propose a novel improvement in the generation process of ADT
% values by exploiting some static information found in the codebase.
%
Then, we provide a compositional approach to capture this information into our
automatically derived random generators.
%
More specifically, we propose a generation process that is capable to consider
how programs branch based on ADTs values' patterns as well as how they get
manipulated by abstract interfaces (Section \ref{sec:hrepcont}).
%
Furthermore, we show how to predict the distribution of the \emph{expected}
numbers of ADT constructors, values fitting certain branching patterns, and
calls to interfaces.
%
For that, we extend some recent results on applying \emph{branching
  processes}\cite{gw1875}---a simple stochastic model conceived to study
population growth (Section \ref{sec:synthesis}).
%
%---to predict the distribution of QuickCheck
%generators\cite{DBLP:conf/haskell/MistaRH18}.
%
We implement our ideas in a tool called \dragenp\footnote{\dragenp is available
  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable to
automatically synthesize \quickcheck generators for \emph{rich ADT values},
where the distributions of random values can be adjusted at compile-time to what
developers might want based on our predictions.
%
Finally, we provide empirical evaluations showing that including static
information from the user codebase improves the code coverage of two external
applications when tested using random values generated following our ideas
(Section \ref{sec:casestudies}).

% \todo[inline,author=Ale]{Add here later about test cases when we know what they
%   are}

We remark that, although this work focuses on Haskell algebraic data types, this
technique is general enough to be applied to most programming languages.
%
%with some level of support for composite types.

### 2019-01-30 paper/randomtesting.tex
General pass. Paper still needs compressing

In this section, we briefly introduce the common approach for automatically
deriving random data generators for ADTs in QuickCheck by using the information
found at the type-level.
%type-driven approach, along
%with its main drawbacks.
%
% Haskell is a strongly typed programming language with a powerful type system.
% %
% It lets programmers encode a considerable amount of information about the
% structure of their systems using data types that can be checked at compilation
% time.
% %
% One of its key aspects is the support for Algebraic Data Types (ADTs).
% %
% Essentially, an ADTs is a composite type defined by combining other types in
% terms of \textbf{sums} (also known as \emph{variant types}) and
% \textbf{products} (or tuples) of other data types.
% %
To exemplify this, and for illustrative purposes, let us consider the following
ADT definition to encode simple |Html| pages:
%
\begin{code}
data Html  =   Text    String
           pp  Single  String
           pp  Tag     String  Html
           pp  Join    Html    Html

\end{code}
%
The type |Html| allows to build pages via four possible constructions:
|Text|---which represents plain text values---, |Single| and |Tag|---which
represent singular and paired Html tags, respectively---, and |Join|---which
concats two Htmls pages one after another.
%
% In the previous definition, we declare |Html| as the sum of four possible
% constructions: |Text| represents plain text values. |Sing| and |Tag| represent
% singular and paired Html tags, and |Join| concats two expression one after
% another.
% %
% We only encode a very small subset of the actual Html specification for
% illustrative reasons.
% %
In Haskell, |Text|, |Single|, |Tag|, and |Join| are known as \emph{data
  constructors} (or constructors for short) and are used to distiguish which
variant of the ADT we are constructing.
%
Each data constructor is defined as a product of zero or more types known as
\emph{fields}.
%
For instance, |Text| has a field of type |String|, whereas |Join| two recusive
fields of type |Html|.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this representation, the page \texttt{<html>hello<hr>bye</html>} can be
encoded as:
%
\begin{code}
Tag "html" (Join (Join
  (Text "hello") (Single "hr")) (Text "bye"))
\end{code}
%
% Additionally, we can define a function |render| to serialize |Html| values as
% follows:
% %
% \begin{code}
% render :: Html -> String
% render (Text t) = t
% render (Sing t) = "<" ++ t ++ ">"
% render (Join x y) = render x ++ render y
% render (Tag t x)
%   = "<" ++ t ++ ">" ++ render x ++ "</" ++ t ++ ">"
% \end{code}

@@

In order to generate random value of types involving ADTs, most approaches
require users to provide a random data generator for each one of them.
%
This is a cumbersome and error prone task and usually \emph{follows closely the ADT
structure}.
%
For instance, consider the following definition of a \quickcheck random
generator for the type |Html|:
%
\begin{code}
genHtml = sized (\size ->
   if size == 0  then  frequency
                       [ (2,  Text    <$> genString)
                       , (1,  Single  <$> genString) ]
   else frequency
     [  (2,  Text    <$> genString)
     ,  (1,  Single  <$> genString)
     ,  (4,  Tag     <$> genString <*> smaller genHtml)
     ,  (3,  Join    <$> smaller genHtml <*> smaller genHtml) ])
\end{code} %$
%
We use the Haskell syntax |[]| and |(,)| for denoting lists and pairs of
elements, respecitvely, e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.
%
The random generator |genHtml| is defined using \quickcheck's function |sized|
to paremetrize the generation process up to an external natural number known as
the \emph{generation size}---capture in the code with variable |size|.
%
This parameter is chosen by the user, and it is used to limit the maximum amount
of recursive calls that this random generator can perform and thus ensuring
the termination of the generation process.
%
When called with a positive generation size, this generator can pick to generate
among any |Html| data constructor of with an explicitly given \emph{generation
  frequency} that can be chosen by the user---in this example, 2, 1, 4 and 3 for
|Text|, |Single|, |Tag|, and |Join|, respectively.
%
When it picks to generate a |Text| or a |Single| data constructor, it also
generates a random |String| value using the standard QuickCheck generator
|genString|.% (\quickcheck provides standard random generators for most base types
% like |String|, |Int|, |Bool|, etc.)
\footnote{%
  The operators |<$>| and |<*>| are used in Haskell to combine values obtained
  from calling random generators and they are not particularly relevant for the
  point being made in this work.}
%$
On the other hand, when it picks to generate a |Join| constructor, it also
generates two independent random subexpression recursively, thus decreasing the
generation size by a unit on each recursive invocation (|smaller genHtml|).
%
The case of random generation of |Tag| constructors follows analogously.
%
This random process keeps calling itself recursively until the generation size
reaches zero, where the generator is constrained to pick only among terminal
data constructors, being |Text| and |Single| the only possible choices in our
particular case.

@@

As many other tools for automatic derivation of generators
(e.g.,\cite{GriecoCB16,RuncimanNL08,mitchell2007,DuregardJW12}), {\dragen}
synthesizes random generators similar to the one shown before, where the
generation process is limited to pick \emph{a single data constructor at the
  time and then recursively generate each required subexpression independently}.
%
% type. the limitations arise quickly when we consider that the underlying
% generation model is essentially the same:
%
% Althogh both \megadeth and \dragen synthesize random generators that are
% theoretically capable to generate the whole space of values of the target data
% type. the limitations arise quickly when we consider that the underlying
% generation model is essentially the same: they pick a single data constructor
% and recusively generate each required subexpression independently.
%
In practice, this procedure is often too generic to generate random data with
enough structural complexity required for testing certain applications.
% to be used for testing purposes.

>>>>

In this section, we briefly introduce the common approach for automatically
deriving random data generators for ADTs in QuickCheck by using the information
found at the type-level.
%
%type-driven approach, along
%with its main drawbacks.
%
% Haskell is a strongly typed programming language with a powerful type system.
% %
% It lets programmers encode a considerable amount of information about the
% structure of their systems using data types that can be checked at compilation
% time.
% %
% One of its key aspects is the support for Algebraic Data Types (ADTs).
% %
% Essentially, an ADTs is a composite type defined by combining other types in
% terms of \textbf{sums} (also known as \emph{variant types}) and
% \textbf{products} (or tuples) of other data types.
% %
%
To exemplify this, and for illustrative purposes, let us consider the following
ADT definition to encode simple |Html| pages:
%
\begin{code}
data Html  =   Text    String
           pp  Single  String
           pp  Tag     String  Html
           pp  Join    Html    Html

\end{code}
%
The type |Html| allows to build pages via four possible constructions:
|Text|---which represents plain text values---, |Single| and |Tag|---which
represent singular and paired Html tags, respectively---, and |Join|---which
concats two Htmls pages one after another.
%
% In the previous definition, we declare |Html| as the sum of four possible
% constructions: |Text| represents plain text values. |Sing| and |Tag| represent
% singular and paired Html tags, and |Join| concats two expression one after
% another.
% %
% We only encode a very small subset of the actual Html specification for
% illustrative reasons.
% %
In Haskell, |Text|, |Single|, |Tag|, and |Join| are known as \emph{data
  constructors} (or constructors for short) and are used to distiguish which
variant of the ADT we are constructing.
%
Each data constructor is defined as a product of zero or more types known as
\emph{fields}.
%
For instance, |Text| has a field of type |String|, whereas |Join| two recusive
fields of type |Html|.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this representation, the example page \texttt{<html>hello<hr>bye</html>}
can be encoded as:
%
\begin{code}
Tag "html" (Join (Join
  (Text "hello") (Single "hr")) (Text "bye"))
\end{code}
%
% Additionally, we can define a function |render| to serialize |Html| values as
% follows:
% %
% \begin{code}
% render :: Html -> String
% render (Text t) = t
% render (Sing t) = "<" ++ t ++ ">"
% render (Join x y) = render x ++ render y
% render (Tag t x)
%   = "<" ++ t ++ ">" ++ render x ++ "</" ++ t ++ ">"
% \end{code}

@@

In order to generate random value of types involving ADTs, most approaches
require users to provide a random data generator for each one of them.
%
This is a cumbersome and error prone task that usually \emph{follows closely the
  ADT structure}.
%
For instance, consider the following definition of a \quickcheck random
generator for the type |Html|:
%
\begin{code}
genHtml = sized (\size ->
  if size == 0
  then frequency
  [  (2,  Text    <$> genString)
  ,  (1,  Single  <$> genString) ]
  else frequency
  [  (2,  Text    <$> genString)
  ,  (1,  Single  <$> genString)
  ,  (4,  Tag     <$> genString <*> smaller genHtml)
  ,  (3,  Join    <$> smaller genHtml <*> smaller genHtml) ])
\end{code} %$
%
We use the Haskell syntax |[]| and |(,)| for denoting lists and pairs of
elements, respecitvely, e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.
%
The random generator |genHtml| is defined using \quickcheck's function |sized|
to paremetrize the generation process up to an external natural number known as
the \emph{generation size}---captured in the code with variable |size|.
%
This parameter is chosen by the user, and it is used to limit the maximum amount
of recursive calls that this random generator can perform and thus ensuring
the termination of the generation process.
%
When called with a positive generation size, this generator can pick to generate
among any |Html| data constructor of with an explicitly given \emph{generation
  frequency} that can be chosen by the user---in this example, 2, 1, 4 and 3 for
|Text|, |Single|, |Tag|, and |Join|, respectively.
%
When it picks to generate a |Text| or a |Single| data constructor, it also
generates a random |String| value using the standard QuickCheck generator
|genString|.% (\quickcheck provides standard random generators for most base types
% like |String|, |Int|, |Bool|, etc.)
\footnote{%
  The operators |<$>| and |<*>| are used in Haskell to combine values obtained
  from calling random generators and they are not particularly relevant for the
  point being made in this work.}
%$
On the other hand, when it picks to generate a |Join| constructor, it also
generates two independent random sub-expression recursively, decreasing the
generation size by a unit on each recursive invocation (|smaller genHtml|).
%
The case of random generation of |Tag| constructors follows analogously.
%
This random process keeps calling itself recursively until the generation size
reaches zero, where the generator is constrained to pick only among terminal
data constructors, being |Text| and |Single| the only possible choices in our
particular case.

@@

As many other tools for automatic derivation of generators
(e.g.,\cite{GriecoCB16,RuncimanNL08,mitchell2007,DuregardJW12}), {\dragen}
synthesizes random generators similar to the one shown before, where the
generation process is limited to pick \emph{a single data constructor at the
  time and then recursively generate each required sub-expression
  independently}.
%
% type. the limitations arise quickly when we consider that the underlying
% generation model is essentially the same:
%
% Althogh both \megadeth and \dragen synthesize random generators that are
% theoretically capable to generate the whole space of values of the target data
% type. the limitations arise quickly when we consider that the underlying
% generation model is essentially the same: they pick a single data constructor
% and recusively generate each required subexpression independently.
%
In practice, this procedure is often too generic to generate random data with
enough structural complexity required for testing certain applications.
% to be used for testing purposes.

### 2019-01-30 paper/relatedwork.tex
General pass. Paper still needs compressing

%
% Boltzmann models
%
Boltzmann models \cite{Duchon2004} are a general approach to randomly generating
combinatorial structures such as trees and graphs, closed simply-typed lambda
terms, etc.
%
A random generator built around such models uniformly generates values of a
target size with a certain size tolerance.
%
However, it has been argued that this approach has theoretical and practical
limitations in the context of software testing \cite{feldt2013}.
%
In a recent work, Bendkowski provides a framework called \emph{boltzmann-brain}
to specify and synthesize standalone Haskell random generators based on
Boltzmann models \cite{bendkowski2018}.
%
This framework mixes parameter tunning and rejection of unwanted samples to
approximate the desired distribution of values according to user demands.
%
In this approach, the overall discard ratio depends on how constrained is this
desired distribution.
%
On the other hand, our work is focused on approximating the desired distribution
as much as possible via parameter optimization, without discarding any generated
value at runtime.
%
Altough promising, we found several difficulties to compare both approaches in
practice due that boltzmann-brain is considered a standalone utility that
produces self-contained samplers.
%
In this light, data specifications have to be manually written using a special
syntax, and cannot include Haskell ground types like |String| or |Int|,
difficulting the task of integrating this tool to existing Haskel codebases like
the ones we evaluate on this work.

%
% GodelTest
%
From the practical point of view, Feldt and Poulding propose G\"odelTest
\cite{feldt2013}, a search-based framework for generating biased data.
%
G\"odelTest uses non-determinism to generate a wide range of data structures.
%
Similar to our approach, it works optimizing the parameters governing the
desired biases on the generated data.
%
However, the optimization mechanism uses metaheuristic search for the best
parameters at runtime.
%
\dragenp on the other hand implements a analytic and composable prediction
mechanism that is used only used at compile time to optimize the generation
parameters, thus avoiding to perform any kind of runtime reinforcement.

%
% DART
%
Directed Automated Software Testing (DART) is a technique that combines random
testing with symbolic execution for C programs \cite{godefroid2005dart}.
%
I requires to instrument the target programs in order to introduce testing
assertions and obtain feedback from previous testing executions that is used to
explore new paths in the source code.
%
This technique has been shown to be remarkably useful, altough it forces a
strong coupling between the testing suite and the target code.
%
Our tool intends to provide better random generation of values following an
undirected fashion, without having to instrument the target code, but still
extracting useful structural information from it.

>>>>

%
% Boltzmann models
%
Boltzmann models \cite{Duchon2004} are a general approach to randomly generating
combinatorial structures such as trees and graphs, closed simply-typed lambda
terms, etc.
%
A random generator built around such models uniformly generates values of a
target size with a certain size tolerance.
%
However, it has been argued that this approach has theoretical and practical
limitations in the context of software testing \cite{feldt2013}.
%
In a recent work, Bendkowski provides a framework called \emph{boltzmann-brain}
to specify and synthesize standalone Haskell random generators based on
Boltzmann models \cite{bendkowski2018}.
%
This framework mixes parameter tunning and rejection of unwanted samples to
approximate the desired distribution of values according to user demands.
%
In this approach, the overall discard ratio depends on how constrained is this
desired distribution.
%
On the other hand, our work is focused on approximating the desired distribution
as much as possible via parameter optimization, without discarding any generated
value at runtime.
%
Altough promising, we found several difficulties to compare both approaches in
practice due that \emph{boltzmann-brain} is considered a standalone utility that
produces self-contained samplers.
%
In this light, data specifications have to be manually written using a special
syntax, and cannot include Haskell ground types like |String| or |Int|,
difficulting the integration of this tool to existing Haskell codebases like the
ones we consider in this work.

%
% GodelTest
%
From the practical point of view, Feldt and Poulding propose \emph{G\"odelTest}
\cite{feldt2013}, a search-based framework for generating biased data.
%
% \emph{G\"odelTest} uses non-determinism to generate a wide range of data structures.
%
Similar to our approach, \emph{G\"odelTest} works by optimizing the parameters
governing the desired biases on the generated data.
%
However, the optimization mechanism uses metaheuristic search to find the best
parameters at runtime.
%
\dragenp on the other hand implements a analytic and composable prediction
mechanism that is used only used at compile time to optimize the generation
parameters, thus avoiding to perform any kind of runtime reinforcement.

%
% DART
%
Directed Automated Software Testing (DART) is a technique that combines random
testing with symbolic execution for C programs \cite{godefroid2005dart}.
%
It requires to instrument the target programs in order to introduce testing
assertions and obtain feedback from previous testing executions, which is used
to explore new paths in the source code.
%
This technique has been shown to be remarkably useful, altough it forces a
strong coupling between the testing suite and the target code.
%
Our tool intends to provide better random generation of values following an
undirected fashion, without having to instrument the target code, but still
extracting useful structural information from it.
